A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-07 paper/paper.Rnw
move brand plots to the part of the paper where we're talking about different results - it's more useful there to explain why performance differs for different classes.

@@

%Labeled images are scarce relative to the amount of data necessary to train a neural network.
\mt{Although transfer learning substantially reduces the amount of data required to train a CNN, l}abeled images are \mt{still difficult or time-consuming to generate} %scarce \fix{[difficult to generate]}
relative to the amount of data necessary to train \mt{even a small} neural network \mt{adequately. Another common} solution to data scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}. Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training; examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

Model training is analogous to the human learning process. For example, a child learns to identify dogs by being presented with many labeled examples, such as when their parent uses the label ``dog" for an animal walking by. That child's understanding of dogs is then measured by how many dogs the child is able to correctly identify, and also by the number of other animals it mistakenly calls ``dog". Similarly, CNN training is a series of stages, or ``epochs", where the model learns from the set of labeled training data, and that learning is measured through intermediate predictions of validation data.

The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since \mt{some categories are much more common than others,}
%the categories do not exist in equal proportion in the labeled data,
the training data were weighted by proportion during the training process to prevent the \mt{model learning} from being overwhelmed by more frequent categories. Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.

\fix{How much should be said about loss and loss functions?}
\mt{A loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions in such a way that will minimize this loss.}

\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process. Overfitting \mt{occurs when a model learns the training data so well that its understanding of the categories becomes specific to only the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.}
%Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
Overfitting is suspected if the validation loss starts to increase after reaching a global minimum. Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated\mt{, because model performance is still improving with continued training}. Neither of these outcomes appears in \autoref{fig:training-accuracy}, indicating that the model optimization process was halted at an appropriate epoch.

\mt{For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to each of the nine shape categories. When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image. Thus, to determine the accuracy of the model predictions, one must choose a threshold to the probabilities, such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no". If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions. Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data. An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.}

\mt{Receiver Operating Characteristic (ROC) curves are a type diagnostic plot that compares the false positive rate against the true positive rate for a classification method, computed by comparing the two rates at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve. Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance.} \autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class. The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.91 (for bowtie and text). While the class performances do vary slightly, each ROC curve is the same general shape and performs \fix{significantly} better than random chance. \mt{The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error. Thus, for each category, any model prediction with probability above the EER for that shape is considered a positive prediction.}

\mt{ROC curves are useful for summarizing the accuracy of model predictions, but they do not indicate which classes are being assigned the false positive or false negative labels when accuracy is not perfect. A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.

In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category. The off-diagonal values represent the proportion of false positives for each shape pairing; \mt{however, since a single image may truly contain multiple shapes}, the values are adjusted to remove the effect of any true positives from the calculation of false positive proportions. For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.

\autoref{fig:confusion_matrix} shows \mt{which classes are being predicted incorrectly, but does not necessarily suggest \emph{why} the confusion is taking place. To investigate further, it is useful to} \fix{see what the model is seeing.}
%the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.

@@

\mt{A heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class. Using these graphics, we can identify which locations in an image contribute most to the model's predictions.} In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.

\fix{I picked the heatmaps I thought illustrated the point best, but didn't modify them heavily yet. How many do we want to include, and do we want to highlight model challenges or just focus on the point that CoNNOR sees what we see?}

<<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
heatmaps <- list.files(file.path(wd, "images", "heatmaps"),
                       pattern = "^heatmap", full.names = T)
knitr::include_graphics(heatmaps[grepl("chevron-circle", heatmaps)])
@

>>>>

%Labeled images are scarce relative to the amount of data necessary to train a neural network.
Although transfer learning substantially reduces the amount of data required to train a CNN \svp{by several orders of magnitude}, \fix{labeled images are still difficult or time-consuming to generate relative to the amount of data necessary to train even a small neural network adequately. Another common solution to data scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}.} \svp{I would lose the last clause of the first sentence and combine it with the next, e.g. to make most efficient use of the labeled data, we enlarge the training data using a process called image augmentation.}
 Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training. Examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

Model training is analogous to the human learning process. For example, a child learns to identify dogs by being presented with many labeled examples, such as when their parent uses the label ``dog" for an animal walking by. That child's understanding of dogs is then measured by how many dogs the child is able to correctly identify, and also by the number of other animals it mistakenly calls ``dog". Similarly, CNN training is a series of stages, or ``epochs", where the model learns from the set of labeled training data, and that learning is measured through intermediate predictions on validation data.

The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since \mt{some categories are much more common than others,}\svp{I like the equal proportion phrasing better myself}
%the categories do not exist in equal proportion in the labeled data,
the training data were weighted by proportion during the training process to \svp{ensure that the model can identify both rare and common geometric shapes}.%from being overwhelmed by more frequent categories.
Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.

\fix{How much should be said about loss and loss functions?}\svp{}
A loss function is a function that measures how far model predictions are from the given labels for each image in the validation and test sets; the model ``learns" by updating its predictions to minimize the loss.

\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process. Overfitting occurs when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.
%Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
\svp{get rid of passive voice here. Overfitting occurs when the validation loss begins to increase after reaching a global minimum} Overfitting is suspected if the validation loss starts to increase after reaching a global minimum. Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated\svp{(for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy})}, because model performance is still improving with continued training. \svp{In \autoref{fig:training-accuracy}, validation accuracy levels off after epoch 20, and validation loss has not yet begun to increase at epoch 30,} indicating that the model optimization process was halted at an appropriate epoch.

For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to \svp{the presence or absence of} each of the nine shape categories \svp{in the image}. When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image. Thus, to determine the accuracy of the model predictions, \svp{fix weird phrasing - we must select a threshold for each label to discretize the model predictions} \fix{one must choose a threshold to the probabilities, such that a predicted probability above the threshold indicates a ``yes" prediction from the model and a probability below the threshold indicates a ``no".} If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions. Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data. An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.

Receiver Operating Characteristic (ROC) curves are a type diagnostic plot \svp{which compare} the false positive rate \svp{to} the true positive rate for a classification method.\svp{Make this two sentences so that it's less convoluted}\fix{, computed by comparing the two rates at a number of thresholds, and the Area Under the Curve (AUC) quantifies the shape of the ROC curve.} Perfect prediction would be indicated by a right angle along the upper left corner of the plot, with a corresponding AUC of 1, and a diagonal line with an AUC of 0.5 would indicate that the classification method performs no better than random chance. \autoref{fig:overall-roc} shows the ROC curve for \fix{the model} across all classes, and \autoref{fig:class-roc} shows the curve for each class. The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for lines) to 0.91 (for bowtie and text). While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance. The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error. \svp{We will use the class-wise EER to set a threshold corresponding to each class, so that }any model prediction with probability above the EER for that shape is considered a positive prediction.

\svp{I'd focus more on the fact that ROC curves are a general summary of the model's performance, but we must also examine the specific cases where the model misclassifies objects to determine if there is a pattern.}
\mt{ROC curves are useful for summarizing the accuracy of model predictions, but they do not indicate which classes are being assigned the false positive or false negative labels when accuracy is not perfect. A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.

In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category. The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values \svp{have been} adjusted to remove the effect of any true positives from the calculation of false positive proportions. \svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}. For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.

\autoref{fig:confusion_matrix} shows \svp{systematic errors in class predictions}, but does not necessarily suggest \emph{why} the confusion is taking place. To assess the reasons for model failure, \svp{we look at the images to} \fix{see what the model is seeing.}
%the general weaknesses of the model, but does not necessarily suggest where to look as to \emph{why} the model is not performing appropriately.

\svp{I would move heatmaps up and out of the diagnostic section and transition into that, possibly}

\svp{Three main model exploration chunks: consistency, heatmaps, brand (motivated by the fact that some classes perform so badly because the exemplars are so variable) - need to figure out the best sequence to transition}

\svp{I've moved this down here because I think it fits better in the model exploration section - using this to explain why performance differs - intraclass variability.}

\paragraph{Brand Investigations}

@@

@@

A heatmap \fix{please add the technical name for heatmaps - look in the distill.io papers?} is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class. Using these graphics, we can identify which locations in an image contribute most to the model's predictions. In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.

\fix{I picked the heatmaps I thought illustrated the point best, but didn't modify them heavily yet. How many do we want to include, and do we want to highlight model challenges or just focus on the point that CoNNOR sees what we see?} \svp{I think it's worth highlighting the fact that we'll never get a model that's perfect because that curly s is pretty much a circle, and there will always be weird shapes that partially activate different categories.}

\svp{We may want to explore using the border of the square to show whether the model actually assigns that label - like the bullet plots}
<<chevron-circle-heatmap, eval = T, out.width = ".95\\textwidth", fig.width = 8, fig.height = 6, fig.scap = "Heatmaps of an image containing chevrons and circles.", fig.cap = "Chevrons are identified by sharp corners and parallel lines, but the thick line between the two shape sections is falsely identified as a quadrilateral.", fig.pos = "!h">>=
heatmaps <- list.files(file.path(wd, "images", "heatmaps"),
                       pattern = "^heatmap", full.names = T)
knitr::include_graphics(heatmaps[grepl("chevron-circle", heatmaps)])
@
